---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
```

# biobench

A curated collection of **real (non-synthetic) biomedical datasets** for
benchmarking classification, regression, and survival machine learning
methods. Every dataset is documented with its canonical literature
reference (`?dataset_name`).

Sourced from CRAN packages only, so every dataset carries clear reuse
terms: [mlbench](https://cran.r-project.org/package=mlbench),
[medicaldata](https://higgi13425.github.io/medicaldata/),
[survival](https://cran.r-project.org/package=survival),
[TH.data](https://cran.r-project.org/package=TH.data),
[KMsurv](https://cran.r-project.org/package=KMsurv),
[timereg](https://cran.r-project.org/package=timereg),
[HSAUR3](https://cran.r-project.org/package=HSAUR3),
[pec](https://cran.r-project.org/package=pec),
[riskRegression](https://cran.r-project.org/package=riskRegression),
[MASS](https://cran.r-project.org/package=MASS),
[boot](https://cran.r-project.org/package=boot),
[aplore3](https://cran.r-project.org/package=aplore3),
[casebase](https://cran.r-project.org/package=casebase),
[addhazard](https://cran.r-project.org/package=addhazard), and
[biostatlab](https://github.com/ielbadisy/biostatlab).

## Installation

```r
# install.packages("remotes")
remotes::install_github("ielbadisy/biobench")
```

## Usage

Datasets load directly via `biobench::<name>`, no `data()` call needed.

```{r example}
library(biobench)

head(breastcancer)

str(gbsg2)
```

## Dataset index

```{r index}
biobench_index()
```

Filter by task, dimensionality, source, size, or missingness:

```{r filter-example}
biobench_index(outcome = "survival", missing = FALSE)
```

## Using biobench in a benchmark study

biobench's job is to supply the datasets; any modeling package can consume
them. The example below uses
[funcml](https://cran.r-project.org/package=funcml), a formula-first ML
framework with a `fit()` / `evaluate()` / `compare()` / `tune()` /
`interpret()` interface, to benchmark several learners on one dataset:

```{r funcml-single, message = FALSE, warning = FALSE}
library(funcml)

d <- na.omit(breastcancer)
d$Id <- NULL # identifier column, not a predictor

compare(
  data = d,
  formula = Class ~ .,
  models = c("glm", "ranger", "xgboost"),
  resampling = cv(5),
  metrics = c("accuracy", "auc"),
  seed = 1
)
```

The same recipe scales to a multi-dataset benchmark study: pick datasets
from `biobench_index()`, pair each with its outcome formula, and loop
`compare()` over them to get one long-format results table.

```{r funcml-loop, message = FALSE}
benchmark_sets <- list(
  breastcancer = list(formula = Class ~ ., data = d),
  pima2        = list(formula = diabetes ~ ., data = na.omit(pima2)),
  dna          = list(formula = Class ~ ., data = dna)
)

results <- lapply(names(benchmark_sets), function(nm) {
  spec <- benchmark_sets[[nm]]
  res <- compare(
    data = spec$data,
    formula = spec$formula,
    models = c("ranger", "xgboost"),
    resampling = cv(5),
    metrics = "accuracy",
    seed = 1
  )
  cbind(dataset = nm, res$results)
})

do.call(rbind, results)[, c("dataset", "model", "mean", "sd", "rank")]
```

`biobench_index(outcome = "classification", missing = FALSE)` (or
`"regression"` / `"survival"`) is the natural way to select which datasets
go into `benchmark_sets` for a larger study.

## Datasets by task

```{r datasets-by-task, echo = FALSE, results = "asis"}
idx <- biobench_index()
idx <- idx[order(idx$task, idx$dataset), ]
db <- tools::Rd_db("biobench")
idx$title <- vapply(idx$dataset, function(nm) {
  rd <- db[[paste0(nm, ".Rd")]]
  tags <- vapply(rd, function(x) attr(x, "Rd_tag"), character(1))
  paste(unlist(rd[[which(tags == "\\title")]]), collapse = "")
}, character(1))

for (tsk in c("classification", "regression", "survival")) {
  sub <- idx[idx$task == tsk, ]
  cat("\n### ", tools::toTitleCase(tsk), " (n = ", nrow(sub), ")\n\n", sep = "")
  cat("| Dataset | Description | n | p | Source |\n|---|---|---:|---:|---|\n")
  for (i in seq_len(nrow(sub))) {
    cat(sprintf(
      "| `%s` | %s | %d | %d | %s |\n",
      sub$dataset[i], sub$title[i], sub$n_row[i], sub$n_col[i],
      sub$source_package[i]
    ))
  }
}
```
