---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
```

# biobench

A curated collection of **real (non-synthetic) biomedical datasets** for
benchmarking classification, regression, and survival machine learning
methods. Every dataset is documented with its canonical literature
reference (`?dataset_name`).

Sourced from CRAN packages only, so every dataset carries clear reuse
terms: [mlbench](https://cran.r-project.org/package=mlbench),
[medicaldata](https://higgi13425.github.io/medicaldata/),
[survival](https://cran.r-project.org/package=survival),
[TH.data](https://cran.r-project.org/package=TH.data),
[KMsurv](https://cran.r-project.org/package=KMsurv),
[timereg](https://cran.r-project.org/package=timereg),
[HSAUR3](https://cran.r-project.org/package=HSAUR3),
[pec](https://cran.r-project.org/package=pec),
[riskRegression](https://cran.r-project.org/package=riskRegression),
[MASS](https://cran.r-project.org/package=MASS),
[boot](https://cran.r-project.org/package=boot),
[aplore3](https://cran.r-project.org/package=aplore3),
[casebase](https://cran.r-project.org/package=casebase),
[addhazard](https://cran.r-project.org/package=addhazard), and
[biostatlab](https://github.com/ielbadisy/biostatlab).

## Installation

```r
# install.packages("remotes")
remotes::install_github("ielbadisy/biobench")
```

## Usage

Datasets load directly via `biobench::<name>`, no `data()` call needed.

```{r example}
library(biobench)

head(breastcancer)

str(gbsg2)
```

## Dataset index

```{r index}
biobench_index()
```

Filter by task, dimensionality, source, size, or missingness:

```{r filter-example}
biobench_index(outcome = "survival", missing = FALSE)
```

## Datasets by task

```{r datasets-by-task, echo = FALSE, results = "asis"}
idx <- biobench_index()
idx <- idx[order(idx$task, idx$dataset), ]
db <- tools::Rd_db("biobench")
idx$title <- vapply(idx$dataset, function(nm) {
  rd <- db[[paste0(nm, ".Rd")]]
  tags <- vapply(rd, function(x) attr(x, "Rd_tag"), character(1))
  paste(unlist(rd[[which(tags == "\\title")]]), collapse = "")
}, character(1))

for (tsk in c("classification", "regression", "survival")) {
  sub <- idx[idx$task == tsk, ]
  cat("\n### ", tools::toTitleCase(tsk), " (n = ", nrow(sub), ")\n\n", sep = "")
  cat("| Dataset | Description | n | p | Source |\n|---|---|---:|---:|---|\n")
  for (i in seq_len(nrow(sub))) {
    cat(sprintf(
      "| `%s` | %s | %d | %d | %s |\n",
      sub$dataset[i], sub$title[i], sub$n_row[i], sub$n_col[i],
      sub$source_package[i]
    ))
  }
}
```
